---
title: "Monty Hall Problem"
author: "Arthur Mello"
date: "1/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Script description

Monty Hall was the original host of a television show called "Let's Make a Deal". The format is quite simple: There are 3 doors. Behind one of them there's a car, and behind 2 of them there's a goat. You have to pick one of the doors and your prize will be whaterver's behind it (obviously you want the car).

Once you have picked a door, the host - Monty Hall - chooses, out of the 2 other doors left, one that has a goat behind it, and opens it, revealing the goat to you. Then, he gives you the choice between sticking to your previously chosen door, or switching. What do you do?

Here is where this problem has raised a lot of furor: our intution tells us there is no difference between changing doors or sticking to our choice, both hold a 50% of hiding the car. A closer look reveals, however, that it is indeed beneficial to change - it will actually double your chances of winning.

Here is a full description of the problem and its solution: https://en.wikipedia.org/wiki/Monty_Hall_problem.

If you don't believe it, however, this script simulates the game 1000 times and show the winning rates for each strategy: changing doors or sticking to the first door. As expected, results are consistent with the calculations.
```{r}
library(rlist)

wins_on_change = 0
wins_on_stay = 0
n = 10000 # n is the number of times we want to simulate the game. It is arbitrary, but it should
        # be high enough to be significant.

for (i in seq(1, n)){
  doors = c(1,2,3)
  
  door_with_prize = sample(x=doors, size=1,
                           replace=FALSE)
  
  # Try1:
  door_i_choose = sample(x=doors, size=1,
                         replace=FALSE)
  doors_left = setdiff(doors, door_i_choose)
  
  if (door_i_choose == door_with_prize){
    
    door_host_chose = sample(x=doors_left, size=1,
                             replace=FALSE)
    
  } else {
    
    door_host_chose = setdiff(doors_left, door_with_prize)
    doors_left = door_with_prize
  }

                           
  # Change:
  door_i_choose_2 = setdiff(doors_left, door_host_chose)
  
  # Answers
  wc = as.integer(door_i_choose_2==door_with_prize)
  ws = as.integer(door_i_choose==door_with_prize)

  wins_on_change = wins_on_change + wc
  wins_on_stay = wins_on_stay + ws                     
  
}

cat("Winning rate on 'switching' strategy: ",
            100*wins_on_change/n,"%", ' \n',
            "Winning rate on 'staying' strategy: ",
            100*wins_on_stay/n,"%", sep = "")

```
